### 2013-09-15 project001/UVM/paper/mcrn501-p1-uvm.abs.tex
working on our paper, integrating paillard model, downloading more data

The explanatory value of simple climate models relies upon their ability to mechanistically explain past variation while providing a reasonable fit to available data.
We investigate the proper fitting of Paillard's three state model with past data to test its accuracy directly.
The parameter fit, our main effort, is achieved via a covariance matrix adaptation evolution strategy (CMA-ES) and a general genetic algorithm (GA).
We find that Paillard's model's explanatory value is mainly empirical, as the model is robust to parameter changes and the fit to data from $\delta ^{18} O$ records has minimum RMSE 1e-5.